The effect of clumpy outflows and disc anisotropy on quasar unification scenarios

### 2015-07-20 paper.tex
added more text

%%%% Various options for document class.
%%\documentclass[usenatbib, a4paper, 11pt]{aastex}
%%\documentclass[preprint, 11pt, a4paper]{aastex}
%%\documentclass[twocolum]{revtex4}
%%\documentclass{report}
\documentclass[useAMS,usenatbib]{mn2e_x}

%\usepackage{psfig,morefloats,url}
%use preprint2 for 2 columns paper.

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
%\usepackage{epsfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

@@

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}

@@

\author[Matthews et al.]{J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge,$^1$
N.~Higginbottom,$^1$ K.~S.~Long,$^2$ S.~A.~Sim,$^3$ and S.~W.~Mangham
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK}

@@

\maketitle

@@

%\begin{keywords}
%AGN: outflows
%\end{keywords}

@@

The spectral energy distributions (SEDs) of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines (e.g. \la, \civ, \nv) 
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to emission from a geometrically thin, 
optically thick accretion disc surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars (REFs). These outflows can take the form of 
highly collimated radio jets \citep{bellonijet2010}(BETTER REF), 
or mass-loaded `winds' emanating from the accretion disc. 
Approximately $20\%$ of quasars exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.

It is possible that the  diverse phenemonology
of luminous AGN and QSOs can be broadly explained by
into a simple picture of geometric unification \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
% As well as acting as a source of photoionized plasma, a 
A disc wind may also have a profound effect on the structure and 
emergent continuum of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
In general, mass-loss treatments appear to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure (REFs), magnetocentrifugal forces \citep{blandfordpayne} and 
radiation pressure on spectral lines \citep[`line-driving'][]{lucysolomon1970, MCGV95}.
Of these, line-driving is possibly the most attractive, due 
to the strong absorption lines seen in BALQSOs and the X-ray spectra of AGN (REFs).
The presence of line-locked features (REFs) and the so-called `ghost of \la' \citep{arav1996}
in the spectra of such objects also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma. \cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.

An alternative solution is that the wind is clumped (REFs), possibly on multiple scale lengths.
Local density enhancements could lower the ionization parameter of the plasma
while still maintaining the same mass-loss rate and column density. 
An inhomogenous outflow has been proposed as a model for the BLR (REFs)
Clumping is expected in outflows and BLRs.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta interacting with either `simple ions'
or `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
The scheme allows one to treat non-LTE line transfer without
approximation for elements which are identified as 
full macro-atoms.

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. To calculate the ionization fractions
of simple-ions the dilute-blackbody modified Saha approach \citep{ML93} is no longer appropriate
due to the presence of a power-law X-ray source. Instead, we model
the SED in a cell using the technique described by H13.

We improve on this method by abandoning the modified Saha approach
entirely, and instead computing the ionization state by 
explicitly solving the rate equations in between ions in non-LTE. 
Photoionization and recombination rates are calculated using MC estimators recorded
during the photon propagation. Not only does this dispense with a number of
small assumptions made in the modified Saha approach, it is also more numerically stable, 
and in principle allows the direct addition of extra physical processes such as Auger ionization, which would have to be approximated if using the previous technique.

@@

We use the same atomic data as described by LK02 and since
updated by H13 and M15, with the addition of direct ionization data from Dere (YEAR). 
Photoionization cross-sections are from Topbase (REF) and  \cite{vfky}.

@@

\bigskip
\bigskip
\bigskip

@@

Here we describe the results from our model.

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.eps}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $75^\circ$ and $80^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:uvspec}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=0.8\textwidth]{figures/uvspec.eps}
\caption
{
Synthetic spectra at four viewing angles in the clumpy model. Plot would look different probably,
and may show composites, but this would be the main synthetic spectrum plot.
}
\label{fig:uvspec}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/opt.eps}
\caption
{
Figure designed to show optical spectrum and comparison to composite.
}
\label{fig:uvspec}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/alpha_ox_both.eps}
\caption
{
X-ray properties of the H13 and clumped model (text filled 
squares), plotted against monochromatic luminosity 
at 2500\AA. Also plotted are the samples considered by
Saez et al. 2012 on a similar plot; The COMBI-7 AGN sample (ref),
the BQS sample (ref) and the Saez et al. (2012) sample of BALQSOs.
}
\label{fig:xray}
\end{figure*}

@@

The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in Quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{43}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

Present a spectrum of the UV and possible optical too.

@@

@@

\begin{enumerate}
\item We have introduced a first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models are not comparable to those in Quasar composite
spectra. This is due to a fundamental constraint discussed in section ?. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry. 
\smallskip
\item We have examined the effect of GR on our disc SED, using the disc atmosphere
and GR ray-tracing code \agn. While including GR effects
does cause the disc SED to become significantly more isotropic,
the effect is not large enough to produce uniform line to continuum ratios
with viewing angle. We discuss other solutions to this problem in section ?; 
It is possible that a combination of GR and reprocessing by the wind could provide a 
solution, and a number of complicated selection effects may be at work
in the building of the quasar composites.
\end{enumerate}

@@

We would like to thank Omer Blaes, Ivan Hubeny, Shane Davis, 
Michael Crenshaw, Nahum Arav, Daniel Proga,

>>>>

%%%% Various options for document class.
%%\documentclass[usenatbib, a4paper, 11pt]{aastex}
%%\documentclass[preprint, 11pt, a4paper]{aastex}
%%\documentclass[twocolum]{revtex4}
%%\documentclass{report}
%\documentclass[useAMS,usenatbib]{mn2e_x}
\documentclass[preprint, a4paper, 11pt]{aastex}
%\usepackage{psfig,morefloats,url}
%use preprint2 for 2 columns paper.

%% declare any packages used
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pdfpages}
\usepackage{appendix}
\usepackage{subfigure}
\usepackage{url}
%\usepackage{epsfig}
%\usepackage[dvips]{color}
%\usepackage{aabib}

@@

%% Journal definitions
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\apjs{ApJs}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}

@@

%\author[Matthews et al.]
\author{
\parbox[t]{\textwidth}{
J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge$^1$,
N.~Higginbottom$^1$, K.~S.~Long$^2$, S.~A.~Sim$^3$ and S.~W.~Mangham$^1$
}
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
}

@@

@@

\maketitle
%\begin{keywords}
%AGN: outflows
%\end{keywords}

@@

The spectral energy distributions (SEDs) of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to emission from a geometrically thin, 
optically thick accretion disc surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
natural feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required by models of galaxy evolution (REFs) 
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
In addition, ADWs may offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
% As well as acting as a source of photoionized plasma, a 
A disc wind may also have a profound effect on the structure and 
emergent continuum of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
Mass-loss therefore appears to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure \citep{weymann1982, begelman1991}, magnetocentrifugal forces 
\citep{blandfordpayne,pelletier_pudritz} and 
radiation pressure on spectral lines \citep[`line-driving';][]{lucysolomon1970,shlosman1985,MCGV95}.
Of these, line-driving is possibly the most attractive, as
strong absorption lines are already seen in BALQSOs and the X-ray spectra of AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
The presence of line-locked features \citep{bowler2014} 
and the `ghost of \la' (Arav et al. 1996; Arav 1996; North 2006; but see 
also Cottis et al. 2010) \nocite{arav1995, arav1996, north2006,cottis2010}
in the spectra of some BALQSOs also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma. \cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.

An alternative solution is that the wind is clumped (REFs), possibly on multiple scale lengths.
Local density enhancements could lower the ionization parameter of the plasma
while still maintaining the same mass-loss rate and column density. 
Clumping is expected in outflows and the BLR in a number of separate models
with different motivations.

Physically, clumping can be present due to magnetic confinement of 
dense regions (REFs) or the instabilities that are theoretically predicted 
in line-driven winds (REFs).

% An inhomogenous outflow has been proposed as a model for the BLR (REFs)
% Clumping is expected in outflows and BLRs.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta interacting with either `simple ions'
or `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
The scheme allows one to treat non-LTE line transfer without
approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions. In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations (see M15).

@@

Macro-atoms have their ion and level populations derived from
MC rate estimators as described by M15. Previously (LK02, H13, M15),
we adopted a modified Saha approach to calculate the ionization fractions
of simple-ions. We have now improved our code to explicitly solve the 
rate equations in between ions in non-LTE. 
Not only does this
dispense with a number of small assumptions made in the modified Saha approach, 
it is also more numerically stable, 
and in principle allows the direct addition of extra physical 
processes such as Auger ionization, which would have to be approximated 
if using the previous technique.

To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
we model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law 
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp}e^{(-h\nu/kT_{exp})}.
\end{equation}
The photoionization rate from ion $i$ can then be written as

\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{n_i \sum_{band~j=0}^{n}}~{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}},
\end{equation}
while the radiative recombination rate is simply given by 
\begin{equation}
R_{i+1,i}(T_e) = \alpha_{i+1,i} n_{i+1} n_e
\end{equation}
$\alpha_{i+1,i}$ here is the recombination rate coefficient from the ground state
of the upper ion to the ground state of the lower ion- we thus neglect recombination
to and from excited states in this calculation. In addition to these
radiative rates, we include dielectronic recombination and collisional ionization
as additive processes. We assume dilute level populations 
for simple-ions.

% We improve on this method by abandoning the modified Saha approach
% entirely, and instead computing the ionization state by 
% explicitly solving the rate equations in between ions in non-LTE. 
% Photoionization and recombination rates are calculated using MC estimators recorded
% during the photon propagation. Not only does this dispense with a number of
% small assumptions made in the modified Saha approach, it is also more numerically stable, 
% and in principle allows the direct addition of extra physical processes such as Auger ionization, which would have to be approximated if using the previous technique.

@@

We use the same atomic data as described by LK02 and since
updated by H13 and M15, with the addition of direct ionization data from Dere (xxxx). 
Photoionization cross-sections are from \textsc{Topbase} \citep{cunto1993} and  \cite{vfky}.
We use dielectronic recombination from \cite{badnell2006} and free-free 
Gaunt factors from \cite{sutherland1998}.

@@

@@

Here we describe the results from our model. This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
We then evaluated these models qualitavely based on the following
criteria:

\begin{itemize}
\item Does the model maintain the correct ionization state and produce strong BALs?
\item Does strong line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\end{itemize}

We arrived at a model which is one of the most promising,
but also representative of a family of models. Examination 
of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5. 
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/agn-grid-archive}.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.eps}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $75^\circ$ and $80^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:uvspec}
\end{figure*} %fullpage

@@

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{ergs~s^{-1}}$\\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}$ 	&	 $v_{esc}$(f=1) \\ 
$R_v$ 	        &	 $10^{19}$cm \\ 
$\alpha$ 	&	 $1.0$ \\
$f$ 	&	 $0.01$ \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
%  $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters used in the model.}
\label{wind_param}
\end{table}

@@

The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in Quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/uvspec.eps}
\caption
{
Synthetic spectra at four viewing angles in the clumpy model. Plot would look different probably,
and may show composites, but this would be the main synthetic spectrum plot.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/opt.eps}
\caption
{
Figure designed to show optical spectrum and comparison to composite.
}
\label{fig:uvspec}
\end{figure*} %fullpage

Figure 4 shows the synthetic spectrum in the UV from the two models, with the
benchmark model from H13 also shown for comparison. We also show a comparison
to composite quasar and BALQSO spectra in figure 5, with the optical waveband 
included. In the UV, strong BAL profiles are formed in both models. This
highlights the first success of our model: clumping allows the correct ionization state 
to be maintained in the presence of strong X-rays. 
The BAL profiles are saturated, 
and generally show the strongest absorption at the terminal velocity.
This contradicts observed profiles of many BALQSOs (REFs), 
although there are some objects which do show this behaviour (REFs). 
However, we have found that this profile shape is by no means ubiquitous.
Figure 6 shows how the line profile varies with angle for three of the UV lines
(see also section 4.4). In addition, the line profile
shape is strongly dependent on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the line profile shape. Figure ?? shows selection of line profiles 
for different kinematic parameters to demonstrate this. 
The degeneracy between parameters, and lack of available
physical and observational constraints goes someway to showing why individual 
fitting of line profile shapes with kinematic models is such a difficult
exercise.

We find that model A can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. However, the red wing of the BAL profiles 
is far stronger than seen in BALQSO spectra and composites.
This is due to a combination of strong collisionally excited emission
and continuum attenuation. To illustrate the amount of 
continuum attenuation we also show the continuum from a disc-only model.
The line-to-continuum ratios at low inclinations are also
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line equivalent widths 
(EWs) -- this is discussed further in section~5.

To assess the ability of the model to match real 
quasar spectra, we show a comparison to an HST quasar composite in figure 5. 
All the spectra are normalised to the flux at 2000\AA in these plots.
Model A matches the \civ\ line profile well, but the \la\ profile
is a factor of $\sim3$ weaker than in the composite spectrum.
Figure 5 also shows the spectrum into the optical,
showing that the outflow produces signiciant optical emission lines
as well as having a noticeable effect on the continuum shape. This
is particularly true around the Balmer jump at $3646\AA$ and is due to
a prominent wind-formed recombination continuum.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/alpha_ox_both.eps}
\caption
{
X-ray properties of the H13 and clumped model (text filled 
squares), plotted against monochromatic luminosity 
at 2500\AA. Also plotted are the samples considered by
Saez et al. 2012 on a similar plot; The COMBI-7 AGN sample (ref),
the BQS sample (ref) and the Saez et al. (2012) sample of BALQSOs.
}
\label{fig:xray}
\end{figure*} %fullpage

@@

\begin{enumerate}
\item We have introduced a first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We have shown that the degree of ionization stratification
in the model is sufficient that LoBAL line profiles
are seen at a subset of viewing angles. 
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models are not comparable to those in Quasar composite
spectra. This is due to a fundamental constraint discussed in section ?. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry. 
\smallskip
\item We have examined the effect of GR on our disc SED, using the disc atmosphere
and GR ray-tracing code \agn. While including GR effects
does cause the disc SED to become significantly more isotropic,
the effect is not large enough to produce uniform line to continuum ratios
with viewing angle. We discuss other solutions to this problem in section ?; 
It is possible that a combination of GR and reprocessing by the wind could provide a 
solution, and a number of complicated selection effects may be at work
in the building of the quasar composites.
\end{enumerate}

@@

We would like to thank Omer Blaes, Ivan Hubeny, Shane Davis, 
Michael Crenshaw, Nahum Arav, Daniel Proga, Daniel Capellupo, Dirk Grupe etc.
